Writing Trash Bin

### 2024-03-22 paper.qmd
Finished the performance section rewrite

The first issue is that individuals will all have their own individual risk utility function, that is, how much they want to avoid or engage with risks or uncertainty. Therefore decision making experiments have an additional layer of noise that value extraction experiments do not. We cannot be sure if particpants answered differently from the ground truth because a visualisation was difficult to read *or* because the participants risk utility function did not align with the one set by the authors. A possible soltion for this is to provide a utility framework for a particular experiment [@Hullman2016], however this method seems to just introduce more noise. @Cheong2016 introduced a very simple optimum strategy into their experiment on bushfire uncertainty visualisations, where participants should choose to evacuate a house as soon as its likelihood of catching fire is 50% or higher. If the participants *did* correctly identify the optimum strategy and answer accordingly, the "decision making experiment" was actually *just a value extraction experiment*. The decision making aspect ceased to matter, all that mattered was identifying if the probability of a fire was greater than 50%. Interestingly, despite this very obvious and simple tactic to maximize payout from the experiment, it seems like many participants did not adopt it, instead acting much more warily as though they were considering whether or not they would *actually* evacuate in the event of a fire. This means it is likely that introducing a payment scheme or a utility function creates too much mental labour for the participants to answer "correctly", and they would rather stick with their own risk utility function that feels *natural* than go through the mental labour of adopting another. Setting up an experiment that requires participants to do laborious calculations or strenuous mental effort for a small marginal benefit is unlikely to result in sucessful participant responses because it is not in the "spirit" of visualisation.

It is clear that audiences primarily interact with visualisations through their "System 1" brain [@daniel2017thinking]. This aligns with other authors comments that visualisation is primarily about "gists" [Spiegelhalter2017]. In this sense, asking participants a question that requires them to shift to their "System 2" is entirely contrary to *why* we use visualisation. The exact reasons we describe visualisation to be so powerful, that they are able to communicate complicated information fast and efficiently *is a product of the system 1 brain* [@daniel2017thinking], however this means displaying information using visualisations have the same weakness' as system 1 thinking, and visualisation authors need to be aware of this. Asking participants to answer questions that require complicated calculations does not reflect how visualisations are used, and will liekly be ignored by the participants (as seen in many of these decision making tasks). While we agree with @Hullman2016 who suggested that what determines an approripate ground truth is largely a philosophical exercise, we differ in the sense that we do not think it should be done, lest you create a needlessly complicated value extraction task.

@@

The final issue with decision making tasks is that calculating "optimum" choices is not a task we use uncertainty for, it is a task that utilises *risk*. Risk and uncertainty are slightly different, as risk is known probabilities and uncertainty is unknown probabilities, so uncertainty is what you get when you cannot accurately define risk [@Spiegelhalter2017]. Communicating risk allows people to weigh up options and make an optimum decision, uncertainty hedges information to let people know there may not even be an optimum strategy. In this sense, many "uncertainty" decision making experiments are actually just risk decision making experiments.

This leaves us with several issues in decision making experiments. They are noisy because of participants individual utility functions, and when that is not the case they are thinly veiled and over complicated value extraction experiments, and on top of that, they dont even evaluate uncertainty as uncertainty, they evaluate it as risk. For these reasons, uncertainty visualisation experiments should either be designed with these caveates in mind and be designed to untangle why a particular visualisation will elicit a specific decision, or the field should steer away using decision quality and related metrics as a basis for evaluating visualisation performance.

#### Trust and confidence
Authors also occasionally use trust as a proxy for uncertainty. Similar to measuring uncertainty through decision quality, this method also has its own series of issues. If the purpose of displaying uncertainty information is to appropriately hedge a signal with noise, then it should be assumed that trust is only related to uncertainty communication through increased transparency and honesty. It does seem to be the case that uncertainty communication increases trust because it is a proxy for transparency *(add citations)*. Unfortunately a large proportion of the literature discusses trust as something that is directly related to uncertainty visualisation rather than as an observable by product. In viewing uncertainty communication as directly related to trust, not related through the proxy of transparency, several unobserved variables are conflated. The amount of uncertainty that is present in the information, whether or not the information is trustworthy or makes sense, prior beliefs of the participants, and the trustworthiness of the source are a few examples of variables that are conflated when authors directly consider trust to be of direct interest. Considering trust, and not transparency, as the metric of importance in uncertainty communication leads to a questionable subtext that argues against transparency. @Hullman2020a found that author simultaneously argued that failing to visualise uncertainty was akin to fraud, but also many avoided uncertainty visualisation because they didn't want their work to come across as "untrustworthy".Being concerned about audiences perception of trust, without first establishing if what we are communicating is *trustworthy*, leads to the implication that details that result in information not being considered trustworthy should be avoided. Authors that imply that a decrease in trust or an increase in ambiguity or risk aversion are metrics of importance in of themselves do not understand the purpose of uncertainty communication as methods to increase transparency. Science communication should be primarily concerned with accuracy, setting trust and risk-aversion as the variables of interest implicitly encourages statisticians to set trust and risk-aversion as the primary goals of communication. The issue of trust being divorced from trustworthiness has been commented on by other authors [@ONeill2018], however the issue still persists in the uncertainty visualisation literature [@Zhao2023].

Additionally, some studies use a ground truth in these trust experiments, which makes for a very confusing read on exactly how participants are supposed to use the information. 
- @Zhao2023 displayed a model prediction with uncertainty and took participants using the model prediction as a sign of trust. They reported that visualising uncertainty information caused participants to trust the model in the low variance case, but the results in the high variance case were inconclusive. The discussion made it clear the authors thought the uncertainty information should make the visualisation more trustworthy, but conflating trust and the use of a prediction implied uncertaitny information should somehow influence participants to use their own prediction. If the Beuro of Meterology was uncertain about their rain prediction, that would not be sufficient conditions for me to decide to make my own rain predictions. Despite this, the authors seemed to assume that the uncertainty information *should* have an influence on that, showing they had not deeply considered how uncertainty inforormation should infludence the choices of the participants.

@@

#### Deterministic questions
- You cannot make a deterministic statements about a random variable (citation: the entire field of probability). Many of these issues just boil down to a really pervasive misunderstanding of statistics. I do not expect the general population to understand this fact about random variables, however *I would expect* researchers investigating *methods of statisticl communication* would at least understand enough to ask appropriate questions that *are answerable*. 
- while we are here, you cant answer a probability question with a deterministic. I cannot say yes or no to a probability question without providing a threshold. 
- Basically, you need a probability threshold to answer these questions, something it does not seem like the authors are collecting, and I am unsure if the authors are aware of this themselves because they never mention it.

#### Other (questionable) attemps to capture uncertainty
There are other studies that try to include uncertainty information without directly asking for it but do not go down the route of asking for a decision or levels of trust. These papers often try to ask a question that should utilise both the uncertainty and signal in the response. This is an interesting approach, but these papers are often still looking for a "signal" which is specified in the form of a "ground truth" that is used to evaluate the plot. Unfortunately this usually comes in the form of slightly cryptic questions that create a large amount of noise on the interpretation side [@Hullman2016].

Cite gap __: Weird questions to capture uncertainty
- @Correll2018 “place your 5 ships on the safest locations on the board.” (minimise danger)
- @Ibrekk1987 asked participants for the "best" estimate while displaying a population density function. The participants proided the mode instead of the mean, although there was no indication the mean was what was being asked about. 
- @Hofmann2012 showed two distributions in 20 different visualisations (a lineup protocol) using a jittered sample, a density plot, a histogram, and a box plot. Participants were asked to report in which of the plots was "the blue group furthest to the right" and to provide reasoning from a multiple choice list and grade their certainty. The experiment set up is shown in @fig-right. The participants answers were then compared to a ground truth where the correct plot had a right shifted mean. By comparing the results to a ground truth statistic and marking participants as "wrong" or "right", the error from the participants that had an alternative interpretation to the concept of "furthest right" was conflated with the error from a the visualisation choice.  
- @Padilla2017 made several confusing assumption when determining the ground truth of their experiments that were assessing storm visualisations. Experiment 2 assumed that the participants should provide responses that indicated they perceived the intensity of the storm to be independent of the uncertainty distribution of its location. This assumption seems counterintuitive as best, however experiment 3 asked participants to "decide which oil rig will receive more damage based on the depicted forecast of the hurricane path" for which they *were* supposed to incorperate uncertainty information and not assume they were independent. The authors seemed to be unaware of these confounding mistakes in their assumptions of how the participants were supposed to utilise the uncertainty information.  
- @Sanyal2009 mapped uncertainty to dots and signal to a 3D surface and asked participants to identify areas of high and low signal and high and low uncertainty. Participants were not asked to combine that information in any way, and the signal and the noise were treated as separate variables.  
- @Correll2014 asked participants to extract the mean and variance from four uncertainty visualisations, bar charts with error bars, box plots, gradient plots, and violin plots (that were adjusted to make sure the mathematical information they were showing was the same) from two side by side distributions. Participants were also asked to answer a question that was supposed to incorperate both the signal and the noise, such as "How likely is candidate B to win the election?" when the two distributions indicated voter preference. Participants were not able to answer the question about likelihood in term of probability, but were instead given seven options from 1=Outcome will be most in favor of A to 7=Outcome will be most in favor of B. The ground truth statistic for this question was a scalar multiple of Cohen’s d, indicating participants were supposed to incorperate uncertainty information using a very specific formula that was likely unknown to them but assumed to be used implicitly. 
- @Cheong2016 tested multiple different visual representations of uncertainty for representing the likelihood of a house being burned down based on its location. Their payment scheme, which paid out $0.10 for a correct choice (i.e. staying when the house was not burned down or leaving when the house was burned down) and 0 for an incorrect choice (i.e.leaving when the house didnt burn down or staying when the house burned down), meant participants were incentivised to base their entire leave/stay decision on whether or not the likelihood of a fire at their house is above or below 50%. The authors failled to recognise that this "complicated decision making task" boilled down to a simple value extraction problem, which the text made easiest to extract. The results from the experiment indicated that many of the participants also failled to recognise the obvious and simply strategy.

- @Blenkinsop2000 tailored their questions according to the visualisation shown, which implies an understanding that different visualisations contain different information. Despite this the authors still compared the visualisations according to which were "useful for uncertainty visualisation".

- @Blenkinsop2000 did try to include uncertainty as noise, asking participants to search for a specific outcome (the land classified as grass) in a random . However the participants failing to identify the signal (something that would be expected of an uncertainty plot) was seen as a sign the task was too complicated. While the task in this experiment *was* certainty far too complicated (and poorly communicated) this discussion indicates the authors did not consider the signal suppression to be the goal of the visualisation, and rather an annoying distraction.

- @Ibrekk1987 asked participants for the "best estimate" which was provided in terms of most probable value (mode), the mean and the median, the "ground truth" statistic used was the mean, despite the "best estimate" largely depending on the loss function of the model and the methods used.

- @Ibrekk1987 used several displays of a probability density functions, since “formally equivalent representations are often not psychologically equivalent”.

- @Ibrekk1987 highlights that in the face of a vague question, participants will use the plot to decide what the authors mean. The authors take this as a

- @Boone2018 organised responses from participants by the approach they used and tried to account for participants prior beliefs (although the assessment of those beliefs used the word "likely" which is open to interpretation).

There is so little understanding of information in uncertainty graphics that participants are frequently provided with information that is insufficient to answer the question at all, an issue that is regularly missed by the experimenters. In order to test whether or not people incorrectly interpreted the cone of uncertainty that is used to communicate cone tasks, experimenters regularly do not give participants enough information to answer the questions [@Padilla2017; @Boone2018]. The cone of uncertainty provides a 95% confidence interval for the *eye* of a hurricane, which allows us to know *where the eye of the storm will likely go*, this does not, give us enough information to answer if the storm will *hit* a particular point, because this requires information about the size and intensity of the storm, *and* an assumption about whether or not the size and intensity will change over time. Since these experiments are often trying to *test* for common misreadings of the plot, such as only thinking the area inside the cone will be damaged or believing the cone represents size and the storm is getting bigger. The authors often fail to recognize that the exact information that has been witheld for testing is necessary for answering their questions, and considering that information is necessary to mark participants as correct or incorrect, it is hard to understand exactly how they evaluated the experiment participants. @Ibrekk1987 expected participants to calculate the mean using a pie chart that had numerical bins combined (e.g. segment one was 0-2 inches of snow, and segment two was 2-4 inches of snow) so it could not be done by calculation. The authors also expected participants to use a CDF to calculate the mean, which needs to be converted *back* into a PDF to do so mathematically, so it was unclear how participants were supposed to do this calculation visually unless they were to just make a blind guess. The authors interpreted this as the participants misinterpreting how the CDF works, not seeming to realise they gave them an impossible task.
In light of this, the main take away from these papers seems to be "if you do not give people the necessary information to answer a question, they will be unable to answer it", which does not seem interesting enough to justify several papers.

@@

@@

## Suggestions to measure uncertainty
While uncertainty visualisation can make it easier to communicate the complicated details of uncertainty information, it carries some challenges that are unique to the visualisaition, specifically with respect to the "signal supression" concept. In simple estimates or verbal communication, the signal is often easy to identify because it is what we are explicitly saying. Visualisations are used in both data exploration and communication. This means what exactly is a *signal* in any particular visualisation is hard to identify, since we often let the visualisation *tell us* what the signal is. Additionally, you cannot add noise to *every single possible* signal one might take from a visualisation. Two people looking at the same visualisation might, just by chance, develop two entirely different insights. These unique and facinating challenges that are faced by viewing uncertainty visualisaiton through the lense of "signal supression" have been almost completely untouched by the literature.

@@

### Just noticiable signal
Connecting our evaluation of uncertainty back to the reasons uncertainty is believed to be different is likely the key to improving evaluation studies. Uncertainty acts as a form of statistical "hedging" for signals found in an analysis. Interviews with experts in statistic back up this primary motivation, as ignoring uncertainty information of often expressed as being similar to fraud or lying and the goal of "improving decisions" is only seen as a secondary outcome of an appropriately hedged signal [@Hullman2020a; @Manski2020].

@@

**Experiment rant from confirmation**

@@

# Great Examples

>>>>

The first issue is that individuals will all have their own individual risk utility function, that is, how much they want to avoid or engage with risks or uncertainty. Therefore decision making experiments that have an additional layer of noise that value extraction experiments do not. We cannot be sure if particpants answered differently from the ground truth because a visualisation was difficult to read *or* because the participants risk utility function did not align with the one set by the authors. Several authors have offered solutions to this issue, however the problem is deeper than any of them seem to realise.

@Hullman2016 suggested providing a utility framework for each experiment, to instruct the participants in how to account for the uncertainty information. This is a method that seems to have been adopted by @Fernandes2018 who describe the following scenario to participants who are trying to maximise the coins in their experiment:
> Subjects gain coins for every minute they are able to continue an activity that is valuable to them (e.g., watching TV at home) before going to the bus stop, and gain a bonus for arriving at their intended destination early. Subjects incur a coin penalty for time spent waiting at a stop for a bus to arrive.

Any payment scheme that is incorporated into an experiment will also implicitly set up a utility framework, as it is assumed participants will try to maximise their payments. In tasks such as this, the ground truth of each question is typically selected to be the value a ration agent would select and visualisations are evaluated on the basis of how far the participants repsponses are from that ground truth. The problem with this, is that all the complexity of the question is in *working out* the best response, and it has little to do with the visualisation itself. The visualisation aspect of these studies becomes a value extraction experiment.

@Cheong2016 tested multiple different visual representations of uncertainty for representing the likelihood of a house being burned down based on its location. Their payment scheme, which paid out $0.10 for a correct choice (i.e. staying when the house was not burned down or leaving when the house was burned down) and 0 for an incorrect choice (i.e.leaving when the house didnt burn down or staying when the house burned down), meant participants were incentivised to base their entire leave/stay decision on whether or not the likelihood of a fire at their house is above or below 50%. If the participants *did* correctly identify the optimum strategy and answer accordingly, the "decision making experiment" was actually *just a value extraction experiment*. The decision making aspect ceased to matter, all that mattered was identifying if the probability of a fire was greater than 50%. Interestingly, despite this very obvious and simple tactic to maximize payout from the experiment, it seems like many participants did not adopt it, instead acting much more warily as though they were considering whether or not they would *actually* evacuate in the event of a fire. This means it is likely that introducing a payment scheme or a utility function creates too much mental labour for the participants to answer "correctly", and they would rather stick with their own risk utility function that feels *natural* than go through the mental labour of adopting another. Setting up an experiment that requires participants to do laborious calculations or strenuous mental effort for a small marginal benefit is unlikely to result in successful participant responses because it is not in the "spirit" of visualisation. (*Cite Gap 16*)

It is clear that audiences primarily interact with visualisations through their "System 1" brain [@daniel2017thinking]. This aligns with other authors comments that visualisation is primarily about "gists" [Spiegelhalter2017]. In this sense, asking participants a question that requires them to shift to their "System 2" is entirely contrary to *why* we use visualisation. The exact reasons we describe visualisation to be so powerful, that they are able to communicate complicated information fast and efficiently *is a product of the system 1 brain* [@daniel2017thinking], however this means displaying information using visualisations have the same weakness' as system 1 thinking, and visualisation authors need to be aware of this. Asking participants to answer questions that require complicated calculations does not reflect how visualisations are used, and will liekly be ignored by the participants. While we agree with @Hullman2016 who suggested that what determines an approripate ground truth is largely a philosophical exercise, we differ in the sense that we do not think it should be done, lest you create a needlessly complicated value extraction task.

@@

The final issue with decision making tasks is that calculating "optimum" choices is not a task we use uncertainty for, it is a task that utilises *risk*. Risk and uncertainty are slightly different, as risk is known probabilities and uncertainty is unknown probabilities, so uncertainty is what you get when you cannot accurately define risk [@Spiegelhalter2017]. Communicating risk allows people to weigh up options and make an optimum decision, uncertainty hedges information to let people know there may not even be an optimum strategy. In this sense, many "uncertainty" decision making experiments are actually just risk decision making experiments. Not only this, but the inclusion of uncertainty information should *not* impact the choices of a rational agent. A well behaved rational angent *should* ignore uncertainty information, so long as the estimate provided is still the best case. This point seems to be lost on many uncertainty visualisation authors. @Zhao2023 displayed a model's prediction and its estimated uncertainty and asked participants if they wanted to submit the model estimate as their answer, or make their own prediction. At no point did the authors seem to realise a rational agent would accept the model estimate every time, no matter how uncertain the information was, and participants were actually disqualified from the study for taking this approach to answering the questions. (*Cite Gap 17*)

This leaves us with several issues in decision making experiments. They are noisy because of participants individual utility functions and when that is not the case they are thinly veiled and over complicated value extraction experiments. Additionally, authors dont even seem to agree how participants should incoreperate uncertainty information, with some authors designing complicated utility functions, and others not realising the best response to their experiment is to ignore uncertainty information entirely. For these reasons, uncertainty visualisation experiments should either be designed with these caveates in mind and be designed to untangle why a particular visualisation will elicit a specific decision, or the field should steer away using decision quality and related metrics as a basis for evaluating visualisation performance.

#### Trust and confidence
Authors also often measure the impact of visualising uncertainty on participants trust in an estimate. Trust is a by product of displaying uncertainty rather than the goal of it, and viewing the relationship in the converse direction can lead to misguided research.

If the purpose of displaying uncertainty information is to appropriately hedge a signal with noise, then it should be assumed that trust is only related to uncertainty communication through increased transparency and honesty. Considering trust, and not transparency, as the metric of importance in uncertainty communication can lead to a questionable subtext that argues against transparency, something that has been noticed by several other authors [@Spiegelhalter2017; ONeill2018]. @Hullman2020a found that author simultaneously argued that failing to visualise uncertainty was akin to fraud, but also many avoided uncertainty visualisation because they didn't want their work to come across as "untrustworthy". These authors are optimising *trust* rather than *transparency*, which means they opt to leave out uncertainty information when it does exactly what it is supposed to, decrease certainty in conclusions.

Science communication should be primarily concerned with accuracy. Setting trust and risk-aversion as the variables of interest implicitly encourages statisticians to set trust and risk-aversion as the primary goals of communication. The issue of trust being divorced from trustworthiness has been commented on by other authors [@ONeill2018], however the issue still persists in the uncertainty visualisation literature. @Zhao2023 displayed a several variations of a visualisation of a model prediction and its uncertainty and took participants using the model prediction as a sign of trust. They reported that visualising uncertainty information caused participants to trust the model in the low variance case, but the results in the high variance case were inconclusive. The discussion made it clear the authors thought the uncertainty information should make the visualisation more trustworthy, but conflating trust and the use of a prediction implied uncertaitny information should somehow influence participants to use their own prediction, even though a prediction being uncertainty does not necessarily mean it is incorrect. Despite this, the authors seemed to assume that the uncertainty information *should* have an influence on that, showing they had not deeply considered *how* uncertainty information should influence the choices of the participants. (*Cite Gap 18*)

@@

#### Other (questionable) attemps to capture uncertainty
There is also a swath of studies that are aware a question that boils down to a value extraction experiment, or a question that should be answered by literally ignoring uncertainty information is not what we want when we consider uncertainty visualisations. These papers often try to ask a question that should utilise both the uncertainty and signal in the response, however this is rarely what actually occurs. This method typically results in in cryptic or confusing questions that create a large amount of noise on the interpretation side of the analysis [@Hullman2016].

Some authors opt for asking slightly vague questions that imply a use of uncertainty, but compare it to a ground truth that is very specific. @Ibrekk1987 asked participants for the "best estimate" which was evaluated in accuracy by comparing it to the mean, however the "best estimate" depends on the loss function we are using, and a loss function of minimised error was not implied by the question. @Hofmann2012 showed two distributions in 20 different visualisations (a lineup protocol) using a jittered sample, a density plot, a histogram, and a box plot and asked participants. Participants were asked to report in which of the plots was "the blue group furthest to the right" The experiment set up is shown in @fig-right. The participants answers were then compared to a ground truth where the correct plot had a blue distribtuon with a right shifted mean. By comparing the results to a ground truth statistic and marking participants as "wrong" or "right", the error from the participants that had an alternative interpretation to the concept of "furthest right" was conflated with the error from a the visualisation choice. These papers make it unclear if the participants got the answers wrong because they misunderstood the question or because of something related to the plot. Therefore, this method leads to inconclusive results about the plot design, and is not advised.

@@

Another method used by authors is to ask a deterministic question about a random event. @Padilla2017 provided participants with a visualisation of the cone of uncertainty and asked then to "decide which oil rig will receive more damage based on the depicted forecast of the hurricane path". The cone of uncertainty provides a 60% confidence interval for the location of the eye of a hurricane, which allows us to know the area where the eye of the storm will go, it does not given any information about the intensity of a storm, the size of a storm, or even if a location will be hit. This inclusion of determinism seems to cause the authors to stumble themselves, as they are not consistent with their assumptions. In their first experiment @Padilla2017 indicated the correct answer was to assume that the storm was equally intense no matter how far from the center of the distribution an oil rig was, however answering their third experiment correctly hinged on assuming the intensity of the storm at a particular point (which in this experiment they phrased as damage) *does* change in intensity as you move away from the center of the distribution. Given these conflicting assumptions, it is unclear how the participants were supposed to adjust the probabilistic path information to answer a deterministic question about which oil rig would receive the most damage. The flipside of this is asking participants for a deterministic answer to a probabilistic question. @Correll2014 asked participants "how likely is candidate B to win the election?" when the two distributions indicated voter preference. Participants were not able to answer the question about likelihood in term of probability, but were instead given seven options from 1=Outcome will be most in favor of A to 7=Outcome will be most in favor of B. The ground truth statistic for this question was a scalar multiple of Cohen’s d, indicating participants were supposed to incorperate uncertainty information using a very specific formula that was likely unknown to them but assumed to be used implicitly.

The final method used by authors is to just explicitly ask about uncertainty and signal information separately. @Sanyal2009 mapped uncertainty to dots and signal to a 3D surface and asked participants to identify areas of high and low signal and high and low uncertainty. Participants were not asked to combine that information in any way, and the signal and the noise were treated as separate variables. @Correll2014 asked participants to separately extract the mean and variance from four uncertainty visualisations. These methods explicitly view the uncertainty and signal as two separate variables that should be extracted from a plot, and not two variables that should be interpreted together. Even viewing these questions as a routine check to make sure the signal information isn't impacted by the uncertainty is counter intuitive, because the whole point *of* the uncertainty is to impact the signal information.

These examples are a bit complete mishmash of methods, however they point to a larger issue that goes beyond decision making, trust and confidence experiments. Authors *have no idea* how to evaluate the effects of uncertainty in an uncertainty visualisation.

@@

## Suggestions to measure uncertainty
So, the current methods of measuring or understanding the role of uncertainty in a visualisation is quesitonable at best, however this is not because visualiation authors are missing the mark, but rather uncertainty is *particularly* difficult to express in a visualisation. In simple estimates or verbal communication, the signal is often easy to identify because it is what we are explicitly saying. Unlike statistical models, visualisations are used in both data exploration and communication. This means what exactly is a *signal* in any particular visualisation is hard to identify, since we often let the visualisation *tell us* what the signal is. Additionally, you cannot add noise to *every single possible* signal one might take from a visualisation. Two people looking at the same visualisation might, just by chance, develop two entirely different insights and draw inference on two completely different statistics. These unique and fascinating challenges that are faced by uncertainty visualisaiton have been completely untouched by the literature. This section will cover some interesting research in uncertainty visualisation and suggestions for better ways to measure uncertainty.

@@

### Just noticiable signal
Alternatively visualisation research could shift away from the accuracy concept all together and either allow for open ended responses, or ask questions that are more abstract to try and better replicate the large structural understanding that comes with the insights gained from visualisation.

@@

**Experiment rant from confirmation**
The field could take a more "exploratory" approach to experimentation and ask multiple accuracy questions for multiple visualisations with different visual encodings and try to identify the *trade-off* in the information a particular plot conveys. Data visualisation is commonly utilised as a tool in data exploration, so it is not uncommon for a data analyst to make a plot with only a vague goal and pull out a large number of adjacent observations. This experimental framework could replicate this process.

@@

# Great Examples 
(good decision making example) @Boone2018 organised responses from participants by the approach they used and tried to account for participants prior beliefs (the method of eliciting prior beleifs was, however, questionable).

### 2024-03-22 writingdumpster.qmd
Finished the performance section rewrite

@@

# Moved over notes
- @Blenkinsop2000 asked different questions for each uncertainty visualisation, indicating they had some awareness that the question has a significant impact on whether or not the visualisation should be used.
- @Gschwandtnei2016 When answering “What is the statistical probability (in %) that the interval has already ended at the marked point in time (red line)?” many answers gave the estimated probability to which the marked point in time falls into the interval. This is correct when asking for the probability to which the interval has already started at the marked position. However, in case of the probability to which the interval has already ended at the marked position, this answer is the inverted probability of what we were looking for. They believe this is their own mistake so they so they don’t present these results.

@@

# Main section outtakes
While this is a focus on uncertainty visualisation, these issues in uncertainty communication need to be acknowledged because they bleed into the uncertainty visualisation literature. Uncertainty visualisation are compared on their ability to communicate explicit risk, elicit trust, and prevent risk or ambiguity avoidance, continuing these issues in uncertainty communication at large.

>>>>

Cite Gap 16: Examples of value extraction decision making tasks

Cite Gap 17: Examples of studies where the correct choice was to *ignore* the uncertainty information

Cite Gap 18: Examples of studies where authors measure trust

@@

# Moved over notes

- @Ibrekk1987 expected participants to calculate the mean using a pie chart that had numerical bins combined (e.g. segment one was 0-2 inches of snow, and segment two was 2-4 inches of snow) so it could not be done by calculation. The authors also expected participants to use a CDF to calculate the mean, which needs to be converted *back* into a PDF to do so mathematically, so it was unclear how participants were supposed to do this calculation visually unless they were to just make a blind guess. The authors interpreted this as the participants misinterpreting how the CDF works, not seeming to realise they gave them an impossible task.
- @Blenkinsop2000 tailored their questions according to the visualisation shown, which implies an understanding that different visualisations contain different information. Despite this the authors still compared the visualisations according to which were "useful for uncertainty visualisation". 
- @Blenkinsop2000 asked different questions for each uncertainty visualisation, indicating they had some awareness that the question has a significant impact on whether or not the visualisation should be used.
- @Gschwandtnei2016 When answering “What is the statistical probability (in %) that the interval has already ended at the marked point in time (red line)?” many answers gave the estimated probability to which the marked point in time falls into the interval. This is correct when asking for the probability to which the interval has already started at the marked position. However, in case of the probability to which the interval has already ended at the marked position, this answer is the inverted probability of what we were looking for. They believe this is their own mistake so they so they don’t present these results.
- @Ibrekk1987 used several displays of a probability density functions, since “formally equivalent representations are often not psychologically equivalent”.
- @Ibrekk1987 highlights that in the face of a vague question, participants will use the plot to decide what the authors mean. 

@@

# Main section outtakes
@Hofmann2012 showed two distributions in 20 different visualisations (a lineup protocol) using a jittered sample, a density plot, a histogram, and a box plot and asked participants. Participants were asked to report in which of the plots was "the blue group furthest to the right" and to provide reasoning from a multiple choice list and grade their certainty. The experiment set up is shown in @fig-right. The participants answers were then compared to a ground truth where the correct plot had a right shifted mean. By comparing the results to a ground truth statistic and marking participants as "wrong" or "right", the error from the participants that had an alternative interpretation to the concept of "furthest right" was conflated with the error from a the visualisation choice.

While this is a focus on uncertainty visualisation, these issues in uncertainty communication need to be acknowledged because they bleed into the uncertainty visualisation literature. Uncertainty visualisation are compared on their ability to communicate explicit risk, elicit trust, and prevent risk or ambiguity avoidance, continuing these issues in uncertainty communication at large.
